---
title: "R - Student Analysis"
author: |
  Mihir Rathod 
  
  2505101010078
date: "2026-09-07"
output:
  html_document: default
  pdf_document: default
---

This project is about analysis on students data to get summary, plots/graphs etc. related to the data.

# 1. Data Importing

```{r, fig.width = 12, out.width = "100%"}
# Load knitr to render wide tables beautifully on one line
library(knitr)

student_data <- read.csv("C:/Users/besto/OneDrive/Desktop/r_lab/ppr/students.csv")
kable(student_data)
```

```{r}
kable(head(student_data))
```

```{r}
str(student_data)
```

```{r}
summary(student_data)
```

# 2. Data Preprocessing and Cleaning

```{r}
colSums(is.na(student_data))
```

```{r}
student_data <- unique(student_data)
```

```{r}
student_data$Gender <- as.factor(student_data$Gender)
student_data$Course <- as.factor(student_data$Course)
student_data$Department <- as.factor(student_data$Department)
student_data$CGPA <- as.numeric(student_data$CGPA)
```

```{r}
colSums(is.na(student_data))
```

# 3. Statistical Analysis

## CGPA Statistics

```{r}
mean(student_data$CGPA)
median(student_data$CGPA)
var(student_data$CGPA)
sd(student_data$CGPA)
min(student_data$CGPA)
max(student_data$CGPA)
```

## Gender Distribution

```{r}
gender_count <- table(student_data$Gender)
gender_count
```

The table function counts the number of students in each gender category.

## Course Distribution

```{r}
course_count <- table(student_data$Course)
course_count
```

The course frequency table shows the number of students enrolled in each course.

# 4. Graph and Plot Creation

## Gender Distribution

```{r}
pie(
  gender_count,
  main = "Gender Distribution",
  labels = paste(names(gender_count), gender_count)
)
```

The pie chart shows the distribution of students according to gender. The size of each section represents the number of students in that category.

## Course Distribution

```{r}
barplot(
  course_count,
  main = "Student Distribution by Course",
  xlab = "Course",
  ylab = "Number of Students",
  las = 2
)
```

The bar plot compares the number of students across different courses.

## Student CGPA

```{r, fig.width = 12, fig.height = 7}
# Increase the bottom margin so Student IDs have breathing room
par(mar = c(10, 5, 4, 2) + 0.1)

barplot(
  student_data$CGPA,
  names.arg = student_data$Student_ID,
  main = "Student CGPA",
  xlab = "", 
  ylab = "CGPA",
  las = 2
)

# Push the x-axis label downwards so it doesn't overlap
title(xlab = "Students", mgp = c(7, 1, 0))
```

The bar plot shows the CGPA of individual students.

## CGPA Distribution

```{r}
hist(
  student_data$CGPA,
  main = "Distribution of Student CGPA",
  xlab = "CGPA",
  ylab = "Number of Students"
)
```

The histogram displays the distribution of CGPA values.

# 5. Course-wise CGPA Analysis

```{r}
course_cgpa <- aggregate(
  CGPA ~ Course,
  data = student_data,
  FUN = mean
)

kable(course_cgpa)
```

The aggregate function calculates the average CGPA for each course.

## Average CGPA by Course

```{r}
barplot(
  course_cgpa$CGPA,
  names.arg = course_cgpa$Course,
  main = "Average CGPA by Course",
  xlab = "Course",
  ylab = "Average CGPA",
  las = 2
)
```

The bar plot compares the average CGPA among different courses.

# 6. Findings

```{r}
overall_cgpa <- mean(student_data$CGPA)
overall_cgpa
```
