---
title: "Student Data Analysis"
author: "TraceFork"
date: "2026-09-22"
output:
  html_document:
    toc: true
    toc_depth: 2
    df_print: paged
---

# Student Data Analysis

This report analyzes the student dataset to generate descriptive statistics, identify distribution patterns, and summarize academic outcomes.

## 1. Data Loading

```{r}
student_data <- read.csv("../data/students.csv")
head(student_data)
```

## 2. Data Structure

```{r}
str(student_data)
summary(student_data)
```

## 3. Missing Value Check

```{r}
colSums(is.na(student_data))
```

## 4. Data Cleaning

```{r}
student_data <- unique(student_data)
student_data$Gender <- as.factor(student_data$Gender)
student_data$Course <- as.factor(student_data$Course)
student_data$Department <- as.factor(student_data$Department)
student_data$CGPA <- as.numeric(student_data$CGPA)
```

## 5. Statistical Summary

```{r}
mean(student_data$CGPA, na.rm = TRUE)
median(student_data$CGPA, na.rm = TRUE)
sd(student_data$CGPA, na.rm = TRUE)
min(student_data$CGPA, na.rm = TRUE)
max(student_data$CGPA, na.rm = TRUE)
```

## 6. Distribution Analysis

```{r}
gender_count <- table(student_data$Gender)
gender_count
```

```{r}
course_count <- table(student_data$Course)
course_count
```

```{r}
hist(student_data$CGPA, main = "Distribution of Student CGPA", xlab = "CGPA", ylab = "Number of Students")
```

## 7. Average CGPA by Course

```{r}
course_cgpa <- aggregate(CGPA ~ Course, data = student_data, FUN = mean)
course_cgpa
```

```{r}
barplot(course_cgpa$CGPA, names.arg = course_cgpa$Course, main = "Average CGPA by Course", xlab = "Course", ylab = "Average CGPA", las = 2)
```

## 8. Conclusion

The dataset shows a broad range of academic performance across courses and departments, with CGPA distribution providing useful insight into overall student outcomes.
